A High Efficiency MAC Protocol for WLANs: Providing Fairness in Dense Scenarios

### 2015-11-03 responseLetter.tex

		`\emph{Performing time slot reservation for each transmission is a well known technique for increasing the throughput and mantanining Quality of Service (QoS) in TDMA schemes, like LTE~\cite{canoLTEcoexistence}. Applying the same concept to CSMA networks by modifying DCF's random backoff proceedure provides similar benefits~\cite{HE}. The following are MAC protocols for WLANs, decentralised and capable of attaining greater throughput than CSMA/CA by constructing collision-free schedules using reservation techniques. A survey of collisio-free MAC protocols for WLANs is presented in~\cite{L_MAC}. In this paper we only overview those that are similar to CSMA/ECA.}'\\

		We attempt to increase the background on the subject by further constraining the category of our proposal. CSMA/ECA, as L-MAC and L-ZC are reservation-like protocols implemented for CSMA in WLANs. In the hopes of putting CSMA/ECA in context with L-MAC and L-ZC, we also included a throughput and fairness comparison.

@@

		\begin{itemize}
			\item {\bfseries The major contributions of this paper come from the two extensions, which is probably not enough considering the journal.}
			\item {\bfseries The difference between this submission and the conference/workshop version is not sufficient. The differences are mainly in evaluation part.}
		\end{itemize}

		We extended our contribution by:
			\begin{enumerate}
				\item Testing the performance of CSMA/ECA under channel errors.
				\item Leveraging the issue of ending up with the largest deterministic backoff with Schedule Reset.
				\item Providing a new figure displaying the average throughput of each group of protocols in a mixed network environment.
				\item Implementing CSMA/ECA$_{\text{Hys}}$ with Schedule Reset in real hardware for the first time.
			\end{enumerate}

		\begin{itemize}
			\item {\bfseries Moreover, the Algorithm does not consider decreasing k, which decreases the contribution of the paper and backward compatibility.}
		\end{itemize}

		We presented Schedule Reset, a mechanism that allows CSMA/ECA to reduce its schedule length without increasing the number of collisions in a perfect channel under saturated conditions. Furthermore, we tested its performance under channel errors and in the real hardware implementation.

@@

		\begin{itemize}
			\item {\bfseries The related work section discusses two MACs in a nice detailed way. However, a more comprehensive related work discussion on CSMA/CA is desired.}
		\end{itemize}

		Space limitations prevented the authors from including further details of CSMA/CA. Nevertheless, CSMA/ECA simply modifies the backoff mechanism of CSMA/CA, which is described by Algorithm 1 in the paper. Authors kindly suggest~\cite{bianchi2000performance} as a well-known study on CSMA/CA performance.

		\begin{itemize}
			\item {\bfseries So, when $2 < N < 7$, each node can send at most 1/8 of the slots. Shouldn't the aggregated throughput of N = 3 be around 1.5 times to the aggregate throughput of N = 2? The reviewer has similar question for later figures regarding to throughput.}
		\end{itemize}

		Yes, in theory. Nevertheless, empty slots and other waiting periods (described by (5) in Section \emph{Throughput bounds of CSMA/ECA$_{\text{Hys+FS}}$}) reduce the channel efficiency. These waiting periods of empty slots are defined in the standard, and ensure fairness among nodes.

@@

		\begin{itemize}
			\item {\bfseries It is clear that Fair Share is to solve the uneven partition caused by Hysteresis. However, k never decreases and it only increases, (at least not shown in Algorithm 3.) A mechanism for decreasing k is needed. Otherwise, k is likely to grow to m = 5 and result in fixed m long waiting time. For example, if k grows to 5, which only requires 5 collisions, the node attempts to send 32 packets every $(2 ^ 5 * 16)/2 - 1 = 255$ slots. 5 collisions might not happen in simulations with pure CSMA/ECA\_{Hys+FS} nodes. However, as this paper suggests, backward compatibility is important. A legacy CSMA/CA can easily contribute 5 collisions to other CSMA/ECA\_{Hys+FS} nodes. On the other aspect, node (even CSMA/ECA\_{Hys+FS} nodes) addition might also introduce unexpected collisions that raise k. If k reduction does not exist, it seems to the reviewer that CSMA/ECA\_{Hys+FS} will eventually converge to CSMA/ECA\_{Hys+ArgMax}.  Moreover, a mechanism for reducing k is also helpful for node removal. Finally, clock drift might also contribute to some collisions. Therefore, the algorithm needs a way to decrease k to make it realistic}
		\end{itemize}

		We appreciate your observation. We now introduced Schedule Reset, which aims at leveraging these issues.

		\begin{itemize}
			\item {\bfseries What is the behavior for non-saturated nodes? When their queue is empty, do they reset their r, k ... everything?}
		\end{itemize}

		Yes. When the MAC queue is emptied, CSMA/ECA nodes reset their backoff stage and retransmission counters to zero ($k$ and $r$, respectively). This means that in non-saturated traffic, Schedule Reset is of little use. Figure~\ref{fig:queueSize} shows the average number of packets in the MAC queue for CSMA/CA and CSMA/ECA.

		\begin{figure}[tb]
		\centering
			\includegraphics[width=0.45\linewidth,angle=-90]{figures/unsaturated/queueSize/queueSize-multiplot-TON.eps}
			\caption{Average MAC queue size in non-saturation}
			\label{fig:queueSize}
		\end{figure}

@@

	\subsection{Regarding ``Minor"}
		\begin{itemize}
			\item {\bfseries Another idea comes to the reviewers mind. If k and CW\_min can be adjusted smartly, for sparse scenario, is the following schedule possible?
    					STA 1:  1   t   1   t;
     					STA 2:  t   1   t   1; 
			i.e., k == 0 and CW\_min == 0. (The reviewer might be wrong.)}
		\end{itemize}

		Having fast and accurate information about the number of contenders in WLANs is a challenging task. In fact, failure to accurately estimate this number is one of the drawbacks of protocols like L-MAC and L-ZC.

		To answer the question: yes, it is possible. Nevertheless, as CSMA/ECA is though to be an evolution of CSMA/CA, we do not use those CW$_{\min}$ values in order to comply with the standard. Also, it is important to keep in mind the transmission durations, described in (5).

@@

		\begin{itemize}
			\item {\bfseries The reviewer is interested in the k values in high contention scenarios. If not too difficult, please make figures for k, especially when CSMA/ECA\_{Hys+FS} co-exist with CSMA/CA. Does CSMA/CA nodes cause collision and increases k? If not, why not?}
		\end{itemize}

		Now that Schedule Reset is introduced we acknowledge that collisions with CSMA/CA nodes, channel errors, slots drifts and addition/withdrawals of contenders will increase CSMA/ECA$_{\text{Hys+FS}}$ nodes's schedule length to its maximum value. This produces an increase in the time between successful transmissions, which the node attempts to reduce with Schedule Reset.

		%The effect of channel errors and slot drift is specially significant in real-hardware implementations. Therefore we provide throughput and lost frames for CSMA/ECA$_{\text{Hys}}$ with an aggressive Schedule Reset. Results show CSMA/ECA outperforming CSMA/CA, and more importantly with a very low percentage of lost frames.
%		
%		It is important to highlight that due to firmware limitations Fair Share could not be implemented. The current version of OpenFWWF only supports IEEE 802.11b/g.

\bibliographystyle{IEEEtran}
\bibliography{../ref}
\end{document}
